Individual heterogeneity in the functional topography of the default-mode network in medial prefrontal cortex

### 2019-03-25 Manuscript.Rmd
Many improvements (see list)
- Fixed ROI list
- Updated FV vs Prop plot
- Deleted unnecessary chunk
- Improved meta analysis permutation speed (mclapply + dedicated function)

Cols <- c("aquamarine4", "#D9541A", rgb(190,190,190,100, maxColorValue = 255)) # left, right, interhemisphere

@@

@@

``` {r Meta-analysis permutation, echo=FALSE, message = FALSE, cache = TRUE}
# This version is literally what I had selected originally, plus 25, OFC, 10v, and RSC
Yeo_labels <- c("L_25_ROI",
                "L_OFC_ROI",
                "L_10v_ROI",
                "R_25_ROI",
                "R_OFC_ROI",
                "R_10v_ROI",
                "L_s32_ROI",
                "L_RSC_ROI",
                "R_RSC_ROI",
                "R_8Ad_ROI",
                "R_9p_ROI",
                "R_23d_ROI",
                "R_d23ab_ROI",
                "R_31a_ROI",
                "R_31pv_ROI",
                "R_31pd_ROI",
                "R_7m_ROI",
                "R_v23ab_ROI",
                "R_p24_ROI",
                "R_d32_ROI",
                "R_9m_ROI",
                "R_p32_ROI",
                "R_a24_ROI",
                "R_10r_ROI",
                "R_10d_ROI",
                "L_8Ad_ROI",
                "L_9p_ROI",
                "L_23d_ROI",
                "L_d23ab_RO",
                "L_31a_ROI",
                "L_31pv_ROI",
                "L_31pd_ROI",
                "L_7m_ROI",
                "L_v23ab_ROI",
                "L_p24_ROI",
                "L_d32_ROI",
                "L_9m_ROI",
                "L_p32_ROI",
                "L_a24_ROI",
                "L_10r_ROI",
                "L_10d_ROI",
                "R_s32_ROI",
                "R_9a_ROI",
                "L_p10p_ROI",
                "L_PCV_ROI")

@@

# For each permutation, reassign the study labels, maintaining the original number of studies per condition
# allStudies <- c(studyList_sv, studyList_dmn)
# nPerms <- 100 # reported statistics are computed @ 5k, but it's computationally demanding to do that every time.
# nullDist <- numeric()
# 
# for (i in seq(nPerms)) {
# 
#   # provide feedback only if requested (if TRUE, it will show up in the doc or pdf file)
#   if (feedback) {
#     if(i %in% c(nPerms*0.25, nPerms*0.5, nPerms*0.75, nPerms*0.95)) {print(paste("Completed", i/nPerms, "percent of the permutation..."))}
#   }
#   
#   # Sample all the studies and divide into 2 groups
#   randPerm <- sample(allStudies)
#   tmpSV <- randPerm[seq(length(studyList_sv))]
#   tmpDMN <- randPerm[-seq(length(studyList_sv))]
# 
#   # Count the number of times each region was reported in the literature
#   tmpDF <- data.frame(Parcel = gsub("_","", Yeo_labels),
#                       SV = sapply(Yeo_labels, function(roi) {length(grep(roi, tmpSV))}),
#                       DMN = sapply(Yeo_labels, function(roi) {length(grep(roi, tmpDMN))}))
# 
#   # Perform the prop test for each pair and get the max chi-squared stat
#   nullDist[i] <- max(
#                   apply(tmpDF[,2:3], 1, function(data) {
#                     prop.test(data, c(length(tmpSV), length(tmpDMN)))$statistic
#                     }
#                   ), na.rm = T)
# 
# }
# 
# # Now do the same for the actual data, to see how it compares to the null dist
# tmpDF <- data.frame(Parcel = gsub("_","", Yeo_labels),
#                     SV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}),
#                     DMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}),
#                     propSV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}) / length(studyList_sv),
#                     propDMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}) / length(studyList_dmn))
# 
# # Perform the prop test for each pair and get the max chi-squared stat + pvalue based on the computed null distribution
# tmpDF$Chi_squared <- apply(tmpDF[,2:3], 1, function(data) {
#   round(prop.test(data, c(length(studyList_sv), length(studyList_dmn)))$statistic, digits = 2)
#   })
# tmpDF$Pval <- 1 - ecdf(nullDist)(tmpDF$Chi_squared) # probability of finding a chi-squared stat at least as large as the observed under a permuted null distribution
# 
# # ROIs that survived the threshold based on Chi-squared value (>95th quantile from null distribution -> 8.43)
# significantROIs <- tmpDF %>% filter(Chi_squared >= quantile(nullDist, 0.95))

@@

``` {r ROI computation (overlapping: old), echo=FALSE, message = FALSE}
#TO DO: CHANGE THIS SO IT'S BASED ON STUDYLIST_SV/DMN, COUNTING JUST LIKE IN THE PERMUTATION ABOVE
## ROI definition
# First, load the raw count csv files
valCounts <- read.csv2('decCounts.csv', sep = ",")
dmnCounts <- read.csv2('dmnCounts.csv', sep = ",")

@@

<!---
```{r Method (Figure 1), out.width = "100%", fig.align = "center", fig.cap= "Summary of the partitioning scheme and its evaluation. Each connectome if transformed into a normalized Laplacian matrix, whose eigenvector associated with the second-to-lowest eigenvalue assigns negative or positive values to each vertex. These continuous maps are binarized by sign, and compared within and between subjects using the chance-correcting Adjusted Rand Index (ARI)."}
# Change to inflated eventually
include_graphics("./Images/method_figure.png")
```
-->

@@

``` {r Stability per subject: Fiedler vector vs prop. DMN plot (Fig 5), fig.align="center", fig.width=3, fig.height=3, fig.cap = "Relationship between the magnitude of Fiedler vector values and the proportion of affiliations of a vertex with DMN throughout 37 sliding windows. Grey lines display each subject's association, while the black line shows the fit from a mixed effects logistic regression.", echo=FALSE, message = FALSE}
# code to generate plot
# Fiedler Vector x proportion of times a vertex was affiliated with DMN
# SC_FVvPDMN_plot <- ggplot(data = allSummaries) +
#   geom_smooth(aes(FiedlerVec, slidePropDMN, group = SubjID), color = "gray", fill = "gray", show.legend = F, alpha = 0.3) +
#   #geom_line(aes(FiedlerVec, FittedVals), alpha = 1, size = 3) +
#   geom_line(data = SC_FVvPDMN_fits, aes(FiedlerVector, Fitted_Proportion), lwd = 3) +
#   labs(x = "Fiedler Vector Values", y = "Proportion of DMN Affiliations") +
#   scale_y_continuous(limits = c(0,1)) +
#   theme_classic()

include_graphics("./Images/SW_props.png")

>>>>

Cols <- c("aquamarine4", "#D9541A", rgb(190,190,190,100, maxColorValue = 255))

@@

# produce one permutation instance of the meta-analysis (so it can be fed to mclapply for quicker results)
metaPerm <- function(list1 = studyList_sv, list2 = studyList_dmn, labels = Yeo_labels, perm = 1) {

  # print every 100th permutation just to have a general sense of how far we are
  if ((perm %% 100) == 0) {sprintf("Permutation %s", perm)}

  # combine study lists
  allStudies <- c(studyList_sv, studyList_dmn)

  # Sample all the studies and divide into 2 groups
  randPerm <- sample(allStudies)
  tmpSV <- randPerm[seq(length(list1))]
  tmpDMN <- randPerm[-seq(length(list1))]

  # Count the number of times each region was reported in the literature
  tmpDF <- data.frame(Parcel = gsub("_", "", labels),
                      SV = sapply(labels, function(roi) {length(grep(roi, tmpSV))}),
                      DMN = sapply(labels, function(roi) {length(grep(roi, tmpDMN))}))

  # Perform the prop test for each pair and get the max chi-squared stat
  maxStat <- max(
    apply(tmpDF[, 2:3], 1, function(data) {
      prop.test(data, c(length(tmpSV), length(tmpDMN)))$statistic
    }
    ), na.rm = T)

  return(maxStat)

@@

``` {r Meta-analysis permutation, echo=FALSE, message = FALSE, cache = TRUE}
# This version is literally what I had selected originally, plus 25, OFC, 10v, and RSC
Yeo_labels <- c("L_25_ROI",
              "L_OFC_ROI",
              "L_10v_ROI",
              "R_25_ROI",
              "R_OFC_ROI",
              "R_10v_ROI",
              "L_s32_ROI",
              "L_RSC_ROI",
              "R_RSC_ROI",
              "R_23d_ROI",
              "R_d23ab_ROI",
              "R_31a_ROI",
              "R_31pv_ROI",
              "R_31pd_ROI",
              "R_7m_ROI",
              "R_v23ab_ROI",
              "R_p24_ROI",
              "R_d32_ROI",
              "R_9m_ROI",
              "R_p32_ROI",
              "R_a24_ROI",
              "R_10r_ROI",
              "R_10d_ROI",
              "L_23d_ROI",
              "L_d23ab_RO",
              "L_31a_ROI",
              "L_31pv_ROI",
              "L_31pd_ROI",
              "L_7m_ROI",
              "L_v23ab_ROI",
              "L_p24_ROI",
              "L_d32_ROI",
              "L_9m_ROI",
              "L_p32_ROI",
              "L_a24_ROI",
              "L_10r_ROI",
              "L_10d_ROI",
              "R_s32_ROI",
              "R_9a_ROI",
              "L_PCV_ROI")

@@

# reported statistics are computed @ 5k, but it's computationally demanding to do that every time.
nPerms <- 1000

# For each permutation, reassign the study labels, maintaining the original number of studies per condition
nullDist <- mclapply(seq(nPerms), function(data) {metaPerm()})

# Now do the same for the actual data, to see how it compares to the null dist
tmpDF <- data.frame(Parcel = gsub("_", "", Yeo_labels),
                    SV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}),
                    DMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}),
                    propSV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}) / length(studyList_sv),
                    propDMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}) / length(studyList_dmn))

# Perform the prop test for each pair and get the max chi-squared stat + pvalue based on the computed null distribution
tmpDF$Chi_squared <- apply(tmpDF[, 2:3], 1, function(data) {
  round(prop.test(data, c(length(studyList_sv), length(studyList_dmn)))$statistic, digits = 2)
  })
tmpDF$Pval <- 1 - ecdf(nullDist)(tmpDF$Chi_squared) # probability of finding a chi-squared stat at least as large as the observed under a permuted null distribution

# ROIs that survived the threshold based on Chi-squared value (>95th quantile from null distribution -> 8.43)
significantROIs <- tmpDF %>% filter(Chi_squared >= quantile(nullDist, 0.95))

@@

``` {r ROI computation (overlapping: DELETE), echo=FALSE, message = FALSE}
#TO DO: CHANGE THIS SO IT'S BASED ON STUDYLIST_SV/DMN, COUNTING JUST LIKE IN THE PERMUTATION ABOVE
## ROI definition
# First, load the raw count csv files
valCounts <- read.csv2('decCounts.csv', sep = ",")
dmnCounts <- read.csv2('dmnCounts.csv', sep = ",")

@@

@@

``` {r Stability per subject: Fiedler vector vs prop. DMN plot (Fig 5), fig.align="center", fig.width=3, fig.height=3, fig.cap = "Relationship between the magnitude of Fiedler vector values and the proportion of affiliations of a vertex with DMN throughout 37 sliding windows. Grey lines display each subject's association, while the black line shows the fit from a mixed effects logistic regression. Dashed red lines indicate the mean FV value at which maps were thresholded.", echo=FALSE, message = FALSE}
# code to generate plot
# Fiedler Vector x proportion of times a vertex was affiliated with DMN
SC_FVvPDMN_plot <- ggplot(data = allSummaries) +
  geom_smooth(aes(FiedlerVec, slidePropDMN, group = SubjID), color = "gray", fill = "gray", show.legend = F, alpha = 0.3) +
  geom_line(data = SC_FVvPDMN_fits, aes(FiedlerVector, Fitted_Proportion), lwd = 2) +
  geom_vline(xintercept = c(-0.014, 0.013), linetype = "dashed", color = Cols[2]) +
  labs(x = "Fiedler Vector Values", y = "Proportion of DMN Affiliations") +
  scale_y_continuous(limits = c(0,1)) +
  theme_classic()

SC_FVvPDMN_plot

#include_graphics("./Images/SW_props.png")
